Taking Dyads Seriously

### 2018-01-09 paper/1_intro.tex
intro and model edits

In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}. We focus on three types of interdependencies that can complicate dyadic analyses. First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict. Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior. We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model. The multiplicative effects capture dependencies that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.

>>>>

In this article, we present the Additive and Multiplicative Effects (AME) model, a Bayesian approach for directly modeling relational data to reflect the data generating process that yields interdependencies in dyadic data structures \citep{hoff:2008,minhas:etal:2016:arxiv}. We focus on three types of interdependencies that can complicate dyadic analyses. First, dependencies may arise within a set of dyads if a particular actor is more likely to send or receive actions such as conflict. Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior. We capture these effects, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model. Third order dependencies capture relationships of transitivity, balance, and clusterability between different dyads -- for example, we can only understand why Poland was involved in a dyadic conflict with Iraq in 2003 if we  understand that the United States invaded Iraq in 2003, and we understand that Poland often participates in US led coalitions. The multiplicative effects capture these sorts of dependencies, especially those that result because the specified model has not accounted for some latent set of shared attributes possessed by actors that affect their probability of interacting with one another.

### 2018-01-09 paper/2_model.tex
intro and model edits

To account for the dependencies that are prevalent in dyadic data, we turn to the AME model. The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages. It is flexible and easy to use for analyzing the kind of relational data often found in social science. It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence. \citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework and the latter article details how it contrasts with alternative network-based approaches. Here we provide a brief review and then move to a simulation exercise.

>>>>

To account for the dependencies that are prevalent in dyadic data, we turn to the AME model. The AME approach can be used to conduct inference on cross-sectional and longitudinal networks with binary, ordinal, or continuous linkages. It is flexible and easy to use for analyzing the kind of relational data often found in social science. It accounts for nodal and dyadic dependence patterns, as well as higher-order dependencies such as homophily and stochastic equivalence.\footnote{\citet{hoff:2015:arxiv,minhas:etal:2016:arxiv} provide a detailed introduction to this framework and the latter article details how it contrasts with alternative network-based approaches. Here we provide a brief review and then move to a simulation exercise.}
